---
title: "Basic inferential data analysis"
author: "pierre attey"
date: "Sunday, January 25, 2015"
output: html_document
---
**Background**

  - The goal of this second part of project is to analyze the ToothGrowth data in the R datasets package through the following steps:
``Load the ToothGrowth data and perform some basic exploratory data analyses`` 
``Provide a basic summary of the data.``
``Use confidence intervals and/or hypothesis tests``
  to compare tooth growth by supp and dose. 
  (Only use the techniques from class, even if 
  there's other approaches worth considering)
``State your conclusions and the assumptions needed for your conclusions``

**1. Load the ToothGrowth data and perform some basic exploratory data analyses**
   
     - Using summary statistics to spot problems by using DescTools Packages
     tools for descriptive statistics

```{r}
library(datasets)
data(ToothGrowth)
library(DescTools)
Desc(ToothGrowth)
```
 
 
The variable dose should be a factor, not numeric we have to assign 
it to be a factor

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```

 - look at the dataset variables after conversion
 
 
```{r}
Desc(ToothGrowth)
```
**2. Provide a basic summary of the data**
```{r}
table(ToothGrowth$dose, ToothGrowth$supp)
```


```{r}
library(ggplot2)
ggplot(aes(x=dose, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=dose))
```



```{r}
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp))
```
**3. Use confidence intervals and/or hypothesis tests**

  - control of the differences between groups assuming unequal inter groups variances
  
```{r}
t.test(len ~ supp, data = ToothGrowth)
```

The confidence interval contains zero because of The p-value. This indicates that we can not reject the null hypothesis that the different supplement types have no effect on tooth length

 -  Subsetting three sub-groups as per dose level pairs

```{r}
ToothGrowth.doses1 <- subset (ToothGrowth, dose %in% c(0.5, 1.0)) 
ToothGrowth.doses2 <- subset (ToothGrowth, dose %in% c(0.5, 2.0)) 
ToothGrowth.doses3 <- subset (ToothGrowth, dose %in% c(1.0, 2.0))
```


  -  control of the differences between all subset groups  

```{r}
t.test(len ~ dose, data = ToothGrowth.doses1)
t.test(len ~ dose, data = ToothGrowth.doses2)
t.test(len ~ dose, data = ToothGrowth.doses3)
```

the p-value for all this test is less than 0.05, and the confidence interval does not contain zero. The mean tooth length increases on raising the dose level. This indicates that we can not accept the null hypothesis, and establish that increasing the dose level leads to an increase in tooth length.

**4. State your conclusions and the assumptions needed for your conclusions**

Supplement type has no effect on tooth growth and Inreasing the dose level leads to increased tooth growth.


